---
title: "Introduction to geniusr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Geniusr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `geniusr` package is designed to help you retrieve data from [Genius](https://genius.com/). Most of this data falls into one of two camps:

- Details about **documents** (e.g. artists, songs) hosted on Genius
- Song **lyrics** hosted on Genius

Both of these areas will be introduced momentarily, once API authentication is dealt with.

## Authentication {#auth}

```{r setup}
library(geniusr)
```

`geniusr` data retrieval is typically done via the [Genius API](https://docs.genius.com/). This necessitates a few common steps:

1. Create a Genius API client via the [API Client management page](https://genius.com/api-clients/new)
2. Generate a client access token from your [API clients page](https://genius.com/api-clients)
3. Pass your access token to a `GENIUS_API_TOKEN` environment variable by explicitly calling the `genius_token()` function, or when prompted by other functions

Once you have set the `GENIUS_API_TOKEN` environment variable, it will be active for the rest of the session and automatically called (via `genius_token()`) when using `geniusr` functions for retrieving data.

## Documents {#docs}



## Lyrics {#lyrics}

