---
title: "Introduction to geniusr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Geniusr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `geniusr` package is designed to help you retrieve data from [Genius](https://genius.com/). Most of this data falls into one of two camps:

- *Genius* **Resources** (e.g. artists, songs, albums)
- Song **lyrics** hosted on *Genius*

Functionality in both of these areas will be demonstrated once API authentication is dealt with.

## Authentication {#auth}

```{r setup}
library(geniusr)
```

`geniusr` data retrieval is typically done by making *authenticated* requests to the [Genius API](https://docs.genius.com/). This necessitates a few common steps:

1. Create a Genius API client via the [API Client management page](https://genius.com/api-clients/new)
2. Generate a client access token from your [API clients page](https://genius.com/api-clients)
3. Pass your access token to a `GENIUS_API_TOKEN` environment variable by explicitly calling the `genius_token()` function, or when prompted by other functions

Once you have set the `GENIUS_API_TOKEN` environment variable, it will be active for the rest of the session and automatically called (via `genius_token()`) when using `geniusr` functions to request data.

## Resources {#resources}

The first family of data retrieval functions help with accessing data from various types of Genius **Resources**, explored in turn below.

### Songs

Functions for requesting data on **songs** hosted on Genius take the form `get_song_xxx()`. For example, `get_song()` returns details about the song itself, as well as user annotations that are attached to it, as a `genius_song` object.

For details about the track *Sing About Me, I'm Dying of Thirst* by Kendrick Lamar:

```{r}
get_song(song_id = 90479)
```

### Artists

Functions for requesting data on an **artist** (i.e. the creator of one or more songs hosted on Genius) take the form `get_artist_xxx()`. For example, `get_artist()` returns details about the artist as a `genius_artist` object

For info about the artist Kendrick Lamar:

```{r}
get_artist(artist_id = 1421)
```

### Albums

Functions for requesting data on an **album** take the form `get_album_xxx()`. For example, `get_album()` returns details about the artist as a `genius_album` object.

For info on the album *Good Kid, M.A.A.D City* by Kendrick Lamar:

```{r}
get_album(album_id = 491200)
```

### Search

Functions for **searching** all content hosted on Genius take the form `search_xxx()`. For example, `search_genius()` returns documents matching a search term(s) as a `genius_resource` object.

We can look for documents that match the search term "Compton".

```{r}
search_genius(search_term = "Compton")
```

## Lyrics {#lyrics}

There are also tools to be found in `geniusr` that help with Genius lyrics retrieval. These functions take the form...you guessed it...`get_lyrics_xxx()`. For example, `get_lyrics_id()` finds song lyrics for a given song id, and returns this data as a tidy tibble - each line of the track as an observation (i.e. a row).

To complete the look, here's song lyrics for *Sing About Me, I'm Dying of Thirst*:

```{r}
get_lyrics_id(song_id = 90479)
```
